---
title: "FPL Ten Gameweek Predictive Model"
author: "Ethan Mitten"
date: "10/13/2021"
output: html_document
---

```{r}
#Load in packages
library(dplyr)
library(tidyverse)
```

```{r}
#Load in Beginning Data
FPL_1920 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2019-20/players_raw.csv")
FPL_2021 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2020-21/players_raw.csv")
NewFPL2021 <- read_csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2020-21/players_raw.csv")
```

```{r}
FPL_1920 <- FPL_1920 %>% filter(total_points > 0)
FPL_1920 <- FPL_1920 %>% filter(assists > 0)
FPL_1920 <- FPL_1920 %>% filter(bonus > 0)
FPL_1920 <- FPL_1920 %>% filter(bps > 0)
FPL_1920 <- FPL_1920 %>% filter(goals_scored > 0)
FPL_1920 <- FPL_1920 %>% filter(influence > 0)
FPL_1920 <- FPL_1920 %>% filter(threat > 0)
FPL_1920 <- FPL_1920 %>% filter(minutes > 0)

NewFPL2021 <- NewFPL2021 %>% filter(total_points > 0)
NewFPL2021 <- NewFPL2021 %>% filter(assists > 0)
NewFPL2021 <- NewFPL2021 %>% filter(bonus > 0)
NewFPL2021 <- NewFPL2021 %>% filter(bps > 0)
NewFPL2021 <- NewFPL2021 %>% filter(goals_scored > 0)
NewFPL2021 <- NewFPL2021 %>% filter(influence > 0)
NewFPL2021 <- NewFPL2021 %>% filter(threat > 0)
NewFPL2021 <- NewFPL2021 %>% filter(minutes > 0)
```


```{r}
#Make linear regression model based on best model from other markdown sheet
library(SciViews)
model <- lm(ln(total_points) ~ scale(assists) + scale(bonus) + scale(bps) + scale(goals_scored) + scale(influence) + scale(threat) + scale(minutes) , data = NewFPL2021)
summary(model)
```

```{r}
#Assumption Modeling
tresid1 <- rstudent(model)
par(mfrow = c(1,2))
qqnorm(tresid1, pch=20)
qqline(tresid1, col="blue")
hist(tresid1)
plot(model$fitted.values, tresid1, pch=20)
abline(h=0)
plot(tresid1, type="o", pch=20)
abline(h=0)
```



```{r}
FPL_1920$scale_assists <- scale(FPL_1920$assists)
FPL_1920$scale_bonus <- scale(FPL_1920$bonus)
FPL_1920$scale_bps <- scale(FPL_1920$bps)
FPL_1920$scale_goals_scored <- scale(FPL_1920$goals_scored)
FPL_1920$scale_influence <- scale(FPL_1920$influence)
FPL_1920$scale_threat <- scale(FPL_1920$threat)
FPL_1920$scale_minutes <- scale(FPL_1920$minutes)
FPL_1920$ln_totalpoints <- ln(FPL_1920$total_points)
```

```{r}
New1920 <- FPL_1920 %>% select(c(scale_assists, scale_bonus, scale_bps, scale_goals_scored, scale_influence, scale_threat, scale_minutes, ln_totalpoints, first_name, second_name))
```

```{r}
#Rename columns in dataset
names(New1920)[names(New1920) == "ln_totalpoints"] <- "total_points"
names(New1920)[names(New1920) == "scale_assists"] <- "assists"
names(New1920)[names(New1920) == "scale_bonus"] <- "bonus"
names(New1920)[names(New1920) == "scale_bps"] <- "bps"
names(New1920)[names(New1920) == "scale_goals_scored"] <- "goals_scored"
names(New1920)[names(New1920) == "scale_influence"] <- "influence"
names(New1920)[names(New1920) == "scale_threat"] <- "threat"
names(New1920)[names(New1920) == "scale_minutes"] <- "minutes"
```


```{r}
#Predict values for 2019-20 season
prediction1920 <- predict(model, newdata = New1920)
New1920$prediction <- round(prediction1920,2)
New1920$difference <- round((New1920$prediction - (New1920$total_points)),2)
```


```{r}
#Look at distribution of difference values for 2019-20
ggplot(New1920, aes(x=difference)) +
  geom_boxplot()

mean(New1920$difference)
```

```{r}
FPL_2122 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/FPLCurrentYearPredict.csv")
```

```{r}
FPL_2122 <- FPL_2122 %>% select(-c(clean.sheets, saves.))
```

```{r}
FPL_2122 <- FPL_2122 %>% filter(total.points > 0)
FPL_2122 <- FPL_2122 %>% filter(assists. > 0)
FPL_2122 <- FPL_2122 %>% filter(bonus. > 0)
FPL_2122 <- FPL_2122 %>% filter(bps. > 0)
FPL_2122 <- FPL_2122 %>% filter(goals.scored > 0)
FPL_2122 <- FPL_2122 %>% filter(influence. > 0)
FPL_2122 <- FPL_2122 %>% filter(threat. > 0)
FPL_2122 <- FPL_2122 %>% filter(minutes. > 0)
```

```{r}
FPL_2122$scale_assists <- scale(FPL_2122$assists.)
FPL_2122$scale_bonus <- scale(FPL_2122$bonus.)
FPL_2122$scale_bps <- scale(FPL_2122$bps.)
FPL_2122$scale_goals_scored <- scale(FPL_2122$goals.scored)
FPL_2122$scale_influence <- scale(FPL_2122$influence.)
FPL_2122$scale_threat <- scale(FPL_2122$threat.)
FPL_2122$scale_minutes <- scale(FPL_2122$minutes.)
FPL_2122$ln_totalpoints <- ln(FPL_2122$total.points)
```

```{r}
New2122 <- FPL_2122 %>% select(c(scale_assists, scale_bonus, scale_bps, scale_goals_scored, scale_influence, scale_threat, scale_minutes, ln_totalpoints, name))
```

```{r}
names(New2122)[names(New2122) == "ln_totalpoints"] <- "total_points"
names(New2122)[names(New2122) == "scale_assists"] <- "assists"
names(New2122)[names(New2122) == "scale_bonus"] <- "bonus"
names(New2122)[names(New2122) == "scale_bps"] <- "bps"
names(New2122)[names(New2122) == "scale_goals_scored"] <- "goals_scored"
names(New2122)[names(New2122) == "scale_influence"] <- "influence"
names(New2122)[names(New2122) == "scale_threat"] <- "threat"
names(New2122)[names(New2122) == "scale_minutes"] <- "minutes"
```

```{r}
prediction_2122 <- predict(model, newdata = New2122)
New2122$prediction <- round(prediction_2122,3)
New2122$difference <- round((New2122$prediction - New2122$total_points),3)
```


```{r}
ggplot(New2122, aes(x=difference)) +
  geom_boxplot()

mean(New2122$difference)
```

#Looking at the model I can use the difference column to see who is projected to have more points than they have at the moment. Gabriel Jesus is at the top of the list with surprisingly Michail Antonio. I have overlooked Antonio as I felt he was overhyped, but so far the data is suggesting good things from both of these players. In terms of who has the highest projected points. My team holds five of the top 20 projected players in it. 

```{r}
gw8 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2021-22/gw8.csv")
gw9 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2021-22/gw9.csv")
gw10 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2021-22/gw10.csv")
#gw11 <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/2021-22/gw11.csv")

#Just keep adding gameweeks to keep it going
```

```{r}
#Make a function to quickly select only the columns we want from each dataset
gameweek_column_cleaner <- function(gameweek_variable) {
  gameweek_variable %>% select(c(total_points,assists,bonus,bps,goals_scored,influence,minutes,threat,name)) 
}
```

```{r}
#Put in gameweek

gw8 <- gameweek_column_cleaner(gw8)
gw9 <- gameweek_column_cleaner(gw9)
gw10 <- gameweek_column_cleaner(gw10)
#gw11 <- gameweek_column_cleaner(gw11)
```

```{r}
#Bring in Clean Dataset for 2021-22 Season
Base_GW <- read.csv("~/Desktop/Data Analytics/Fantasy PL Project/FantasyPL_Data/FPLCurrentYearPredict.csv")
Base_GW <- Base_GW %>% select(-c(clean.sheets, saves.))
```

```{r}
#Rename columns in dataset
names(Base_GW)[names(Base_GW) == "total.points"] <- "total_points"
names(Base_GW)[names(Base_GW) == "assists."] <- "assists"
names(Base_GW)[names(Base_GW) == "bonus."] <- "bonus"
names(Base_GW)[names(Base_GW) == "bps."] <- "bps"
names(Base_GW)[names(Base_GW) == "goals.scored"] <- "goals_scored"
names(Base_GW)[names(Base_GW) == "influence."] <- "influence"
names(Base_GW)[names(Base_GW) == "threat."] <- "threat"
names(Base_GW)[names(Base_GW) == "minutes."] <- "minutes"
```


```{r}
#Keep adding to make next projections
gw8_projections <- rbind(Base_GW, gw8)
gw9_projections <- rbind(gw8_projections, gw9)
gw10_projections <- rbind(gw9_projections, gw10)
#gw11_projections <- rbind(gw10_projections, gw11)

#Current Week's Projection Model
X = gw10_projections
```

```{r}
#Take Out 0 values for natural log scaling
X <- X %>% filter(total_points > 0)
X <- X %>% filter(assists > 0)
X <- X %>% filter(bonus > 0)
X <- X %>% filter(bps > 0)
X <- X %>% filter(goals_scored > 0)
X <- X %>% filter(influence > 0)
X <- X %>% filter(threat > 0)
X <- X %>% filter(minutes > 0)
```

```{r}
#Make new columns for scaled variables
X$scale_assists <- scale(X$assists)
X$scale_bonus <- scale(X$bonus)
X$scale_bps <- scale(X$bps)
X$scale_goals_scored <- scale(X$goals_scored)
X$scale_influence <- scale(X$influence)
X$scale_threat <- scale(X$threat)
X$scale_minutes <- scale(X$minutes)
X$ln_totalpoints <- ln(X$total_points)
```

```{r}
#Only Take Scaled Variables for XX dataset
XX <- X %>% select(c(scale_assists, scale_bonus, scale_bps, scale_goals_scored, scale_influence, scale_threat, scale_minutes, ln_totalpoints, name))
```

```{r}
#Rename Columns 
names(XX)[names(XX) == "ln_totalpoints"] <- "total_points"
names(XX)[names(XX) == "scale_assists"] <- "assists"
names(XX)[names(XX) == "scale_bonus"] <- "bonus"
names(XX)[names(XX) == "scale_bps"] <- "bps"
names(XX)[names(XX) == "scale_goals_scored"] <- "goals_scored"
names(XX)[names(XX) == "scale_influence"] <- "influence"
names(XX)[names(XX) == "scale_threat"] <- "threat"
names(XX)[names(XX) == "scale_minutes"] <- "minutes"
```

```{r}
#Model Predictions for Each Current Gameweek
live_predictions <- predict(model, newdata = XX)
XX$prediction <- round(live_predictions,3)
XX$difference <- round((XX$prediction - XX$total_points),3)
```


```{r}
#Graph and Evaluate Each Weeks Performance
ggplot(XX, aes(x=difference)) +
  geom_boxplot()

mean(X$difference)
```

Week 8

#Still has similar people holding the highest predictive differences. Gabriel Jesus, Michail Antonio, Romelu Lukaku...

Week 9

#The model has been changed and is following a scaled transformation for now, so what is seen is that some players that are expected to score more points are:

*Dennis
*Kane 
*McGinn
*Smith-Rowe
*Salah
*Keita
*Bailey
*Richarlison
*Jones
*Werner
*Foden

#We will see how this players trend in the coming weeks and how this list moves

Week 10...

#A lot of movement this week and Jarrod Bowen for West Ham takes a clear number one spot with his residual value being one. Dennis, Kane, McGinn and Smith-Rowe continue to be up there. Ronaldo has also pushed his way into the conversation as well, so we will continue to keep track of these players. More closely we will keep track of Bowen, McGinn and Smith-Rowe as we are most concerned with midfielders.
























